Perturbing SU($n$)-symmetric interactions

### 2020-03-05 collective_sun/sun_perturbation.tex
clarified an association between an operator and its restriction to a subspace
minor change in notation

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the
permutationally symmetric manifold $\M_0$ by the two-body perturbation
$\V_2$ in \eqref{eq:perturbations} is simply the restriction of $\V_2$
onto $\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQPS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQPS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-spin
operator $O_{pq}$.  If the two-spin operator
$O=X\otimes Y+Y\otimes X$, then
\begin{align}
  H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} + \col{Y}\,\col{X} - \col{XY} - \col{YX}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q}
  \sp{\sum_k \p{h_{pk} w_{Okq} + h_{qk} w_{Opk}}
  - \bar h_{pq} w_{Opq}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h} \c \v{\m w} = \Delta \v{\m w},
  \label{eq:V2_eig}
\end{align}
where
\begin{align}
  \check{\m h}
  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
    \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
    - \bar h_{pq} \bra{\set{p,q}}}.
  \label{eq:V2_eig_mat}
\end{align}
We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:V2_eig_mat} because
these terms vanish in \eqref{eq:H_V2_psi} from the fact that
$h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
\eqref{eq:V2_eig} requires diagonalizing the
${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
$\check{\m h}$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
perform this reduction in Appendix \ref{sec:eigenstates_TI}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:V2_eig} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:V2_eig} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold;
$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
$W_k$ are scalar coefficients.  In order to expand the coefficients
$W_k$, we define
\begin{align}
  \v{\m w}=\sum_{p<q} w_{pq} \ket{\set{p,q}},
  &&
  \v w \equiv \sum_{p,q} w_{pq} \ket{p},
\end{align}
where $w_{pq}$ is defined for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{pq}=w_{qp}$ and $w_{pp}=0$.  The vector $\v w$ is the
sum of all columns of the matrix
$\m w\equiv\sum_{p,q}w_{pq}\op{p}{q}$.  We then denote the projection
of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:V2_eig_mat} by $\v{\m w}_\Delta$,
and define
\begin{align}
  w_{\t{columns}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v w_\Delta\c\v w_\Delta}{\Delta},
  &&
  w_{\t{pairs}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v{\m w}_\Delta\c\v{\m w}_\Delta}{\Delta},
\end{align}
in terms of which
\begin{align}
  W_4 \equiv \f{w_{\t{columns}}^{(2)} - w_{\t{pairs}}^{(2)}}{N!_4},
  &&
  W_2 \equiv \f{w_{\t{columns}}^{(2)} - 2 w_{\t{pairs}}^{(2)}}
  {4N!_2} + \f12 \p{3N-4} W_4,
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:H_VM_psi_eig} in a basis that obeys both permutational
symmetry and translational invariance.  To this end, we first expand
the coupling tensor $\m w$ in a complete basis for symmetric tensors:
\begin{align}
  \m w = \sum_{k\in\C_N\p{M}} w_k \m m_k,
  &&
  \m m_k \equiv \sum_{\sigma\in P_M} \m e_{\sigma\p{k}},
  \label{eq:sym_tensor}
\end{align}
where $P_M$ is the set of all permutations of $M$ symbols, and
$\m e_\ell$ is a unit tensor with one in the entry indexed by $\ell$
and zeros everywhere else.  On a $D$-dimensional lattice of linear
size $L$, translational invariance implies that the coefficients $w_k$
remain the same if all spins indexed by $k$ are shifted by the same
displacement $d\in\ZZ_L^D$.  To account for this invariance, we denote
a translation of all spins in $k$ by a displacement $d\in\ZZ_L^D$ by
$k_d$, and define the translation-symmetrized basis tensors
\begin{align}
  \m m_k^{\t{TS}}
  \equiv \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \m m_{k_d}
  = \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \sum_{\sigma\in P_M}
  \m e_{\sigma(k_d)},
\end{align}
where the normalization factor $\N_k^{\t{TS}}$ enforces that the
entries of $\m m_k^{\t{TS}}$ are all either zero or one, which
otherwise might not occur if some translations of $k$ results in the
same choice of spins.  We then define the set $\T_{LD}\p{M}$ of
equivalence classes in $\C_N\p{M}$ under arbitrary translations on
$D$-dimensional lattice of linear size $L$, and expand
\begin{align}
  \m w = \sum_{k\in\T_{LD}\p{M}} w_k \m m_k^{\t{TS}}.
  \label{eq:sym_shift_tensor}
\end{align}
Having expanded $\m w$ in a basis that respects permutational symmetry
and translational invariance, we need to express the eigenvalue
equation \eqref{eq:H_VM_psi_eig} in the same basis.  In order to do
so, we label spins $p,q$ by the position of the corresponding lattice
site in $\ZZ_L^D$, and use translational invariance of the couplings
$\m h$ to expand
\begin{align}
  \m h \equiv \sum_{p,q\in\ZZ_L^D} h_{pq} \op{p}{q}
  = \sum_{q,d\in\ZZ_L^D} h_{q+d,q} \op{q+d}{q}
  = \sum_{d\in\ZZ_L^D} h_{d,0} \m s_d,
  &&
  \m s_d \equiv \sum_{q\in\ZZ_L^D} \op{q+d}{d}.
\end{align}
Starting from this form of $\m h$ and the expansion of $\m w$ in
\eqref{eq:sym_tensor}, we can work out that
\begin{align}
  \m h\circ\m w
  = \sum_{k\in\T_{LD}\p{M}} \sum_{d\in\ZZ_L^D} \sum_{a=1}^M w_{k_{d;a}}
  h_{d,0} w_{k_{d;a}} \m m_k^{\t{TS}},
\end{align}
where $k_{d;a}$ denotes the spins acquired by displacing the $a$-th
spin of $k$ by $d$.  With a bit of additional work, we can also
simplify
\begin{align}
  \bar{\m h}
  = \sum_{k\in\T_{LD}\p{M}} \sp{M h_0 - \sum_{p,q\in k} h_{pq}}
  \m m_k^{\t{TS}},
  &&
  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0},
\end{align}
which implies that
\begin{align}
  \bar{\m h} * \m w = M h_0 \m w
  - \sum_{k\in\T_{LD}\p{M}} \sum_{p,q\in k} h_{pq} w_k \m m_k^{\t{TS}}.
\end{align}
Putting everything together, the generalized eigenvalue problem in
\eqref{eq:H_VM_psi_eig} can be written in the form
\begin{align}
  \sum_{k\in\T_{LD}\p{M}} \sp{\sum_{p,q\in k} h_{pq} w_k
    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}}}
  \m m_k^{\t{TS}}
  = \p{M h_0 + \Delta} \m w,
\end{align}
or equivalently
\begin{align}
  \ut{\check{\m h}} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
  \label{eq:H_VM_eig_simp}
\end{align}
where
\begin{align}
  \ut{\check{\m h}} \equiv
  \sum_{k\in\T_{LD}\p{M}} \ket{k}
  \sp{\sum_{p,q\in k} h_{pq} \bra{k}
    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M \bra{k_{d;a}}},
  &&
  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k},
\end{align}
and, to keep notation simple, we define $\ket{k}=0$ for all $k$ with
repeated values, and enforce $\ket{j}=\ket{k}$ if spins $j$ and spins
$k$ are the same up to a global translation.

@@

Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
can be treated perturbatively, resulting in an effective Hamiltonian
on a permutationally symmetric manifold $\M_0$ of $N$ spins.
Computing this effective Hamiltonian generally requires multiplying
several operators, and projecting the resulting product onto $\M_0$.
Even if the perturbation $\V_M$ is too strong for rigorous
perturbative treatment, as long as it is comparable in magnitude to
the energetic penalty incurred by breaking permutational symmetry, its
effects can still be accounted for by an effective theory on the joint
manifold of permutationally symmetric states together with states of
``weak'' permutational symmetry breaking, e.g.~the manifold of states
generated by the action of $M$-body perturbations on permutationally
symmetric states.  The expansion of operators in a basis for this
joint manifold likewise requires multiplying several operators and
projecting the resulting product onto $\M_0$.  Specifically, if
$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
\begin{align}
  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
\end{align}
where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
entries with a repeated index).  Denoting a complete basis of
permutationally symmetric states by $\set{\ket{m}}$, we can then
identify the $M$-body perturbed states
\begin{align}
  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
  &&
  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
\end{align}
The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
expanded Hilbert space is then
\begin{align}
  \Q \simeq_M \P_M \Q \P_M,
  &&
  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
\end{align}
where finding matrix elements of $\Q$ now requires computing products
of the form
\begin{align}
  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
  \label{eq:example_prod}
\end{align}
The purpose of this section is to work out such products in a
systematic way.  To this end, we derive a general diagrammatic
expansion of these products, as well as rules to simplify the
resulting expansions so as to minimize their computational cost.
These diagrammatic methods allow us to algorithmically evaluate
products such as \eqref{eq:example_prod} from the bare multi-body
operator data $\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out
some special cases of products that are relevant to the perturbation
theory in Sections \ref{sec:single_body_pert} and
\ref{sec:two_body_pert}.

>>>>

The first-order effective Hamiltonian $H_2^{(1)}$ induced on the
permutationally symmetric manifold $\M_0$ by the two-body perturbation
$\V_2$ in \eqref{eq:perturbations} is simply the restriction of $\V_2$
onto $\M_0$:
\begin{align}
  H_2^{(1)} = \sum_O \sum_{p<q} w_{Opq} \P_0 O_{pq} \P_0
  \EQPS \sum_O \sum_{p<q} w_{Opq} O_{1,2}
  \EQPS \sum_O \bar w_O \col{O},
\end{align}
where
\begin{align}
  \bar w_O \equiv {N \choose 2}^{-1} \sum_{p<q} w_{Opq},
  &&
  \col{O} \equiv \sum_{p<q} O_{pq},
\end{align}
are an average of $w_{Opq}$ and a collective version of the two-spin
operator $O_{pq}$.  If the two-spin operator
$O=X\otimes Y+Y\otimes X$, then
\begin{align}
  H_2^{(1)} \EQPS \f12 \sum_{\p{X,Y}} \bar w_{\p{X,Y}}
  \p{\col{X}\,\col{Y} + \col{Y}\,\col{X} - \col{XY} - \col{YX}}.
\end{align}
In order to compute the second-order effective Hamiltonian
$H_2^{(2)}$, as before we pick an arbitrary permutationally symmetric
state $\ket\psi\in\M_0$ and expand (see Appendix
\ref{sec:eigenstates})
\begin{align}
  H_0 \V_2 \ket\psi
  = E_0 \V_2 \ket\psi
  + \sum_O \sum_{p<q}
  \sp{\sum_k \p{h_{pk} w_{Okq} + h_{qk} w_{Opk}}
  - \bar h_{pq} w_{Opq}}
  O_{pq} \ket\psi,
  \label{eq:H_V2_psi}
\end{align}
where to keep notation compact, we define
\begin{align}
  \bar h_{pq} \equiv h_p + h_q - 2 h_{pq},
  &&
  h_p \equiv \sum_q h_{pq},
\end{align}
and we define $w_{Opq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{Opq}=w_{Oqp}$ and $w_{Opp}=0$.  We thus find that the
vector $\V_2^\Delta\ket\psi$ (indexed by $\Delta$) is an eigenvector
of the interaction Hamiltonian $H_0$ if its coefficients
$\v{\m w}_O^\Delta$ all satisfy the eigenvalue equation
\begin{align}
  \check{\m h}_2 \c \v{\m w} = \Delta \v{\m w},
  \label{eq:V2_eig}
\end{align}
where
\begin{align}
  \check{\m h}_2
  \equiv \sum_{p<q} \ket{\set{p,q}} \sp{\sum_{k\notin\set{p,q}}
    \p{h_{pk} \bra{\set{k,q}} + h_{qk} \bra{\set{p,k}}}
    - \bar h_{pq} \bra{\set{p,q}}}.
  \label{eq:V2_eig_mat}
\end{align}
We explicitly exclude $k\in\set{p,q}$ in \eqref{eq:V2_eig_mat} because
these terms vanish in \eqref{eq:H_V2_psi} from the fact that
$h_{kk}=w_{Okk}=0$.  Solving the general eigenvalue problem in
\eqref{eq:V2_eig} requires diagonalizing the
${N \choose 2}\times{N \choose 2}\sim N^2\times N^2$ matrix
$\check{\m h}_2$.  When both he interaction Hamiltonian $H_0$
translationally invariant, however, this eigenvalue problem can be
reduced to that of diagonalizing an $\sim N\times N$ matrix; we
perform this reduction in Appendix \ref{sec:eigenstates_TI}.

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:V2_eig} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:V2_eig} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold;
$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
$W_k$ are scalar coefficients.  In order to expand the coefficients
$W_k$, we define
\begin{align}
  \v{\m w}=\sum_{p<q} w_{pq} \ket{\set{p,q}},
  &&
  \v w \equiv \sum_{p,q} w_{pq} \ket{p},
\end{align}
where $w_{pq}$ is defined for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{pq}=w_{qp}$ and $w_{pp}=0$.  The vector $\v w$ is the
sum of all columns of the matrix
$\m w\equiv\sum_{p,q}w_{pq}\op{p}{q}$.  We then denote the projection
of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}_2$ defined in \eqref{eq:V2_eig_mat} by
$\v{\m w}_\Delta$, and define
\begin{align}
  w_{\t{columns}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v w_\Delta\c\v w_\Delta}{\Delta},
  &&
  w_{\t{pairs}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v{\m w}_\Delta\c\v{\m w}_\Delta}{\Delta},
\end{align}
in terms of which
\begin{align}
  W_4 \equiv \f{w_{\t{columns}}^{(2)} - w_{\t{pairs}}^{(2)}}{N!_4},
  &&
  W_2 \equiv \f{w_{\t{columns}}^{(2)} - 2 w_{\t{pairs}}^{(2)}}
  {4N!_2} + \f12 \p{3N-4} W_4,
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Our task is essentially to express the multi-body eigenvalue problem
in \eqref{eq:H_VM_psi_eig} in a basis that obeys both permutational
symmetry and translational invariance.  To this end, we first expand
the coupling tensor $\m w$ in a complete basis for symmetric tensors:
\begin{align}
  \m w = \sum_{k\in\C_N\p{M}} w_k \m m_k,
  &&
  \m m_k \equiv \sum_{\sigma\in P_M} \m e_{\sigma\p{k}},
  \label{eq:sym_tensor}
\end{align}
where $P_M$ is the set of all permutations of $M$ symbols, and
$\m e_\ell$ is a unit tensor with one in the entry indexed by $\ell$
and zeros everywhere else.  On a $D$-dimensional lattice of linear
size $L$, translational invariance implies that the coefficients $w_k$
remain the same if all spins indexed by $k$ are shifted by the same
displacement $d\in\ZZ_L^D$.  To account for this invariance, we denote
a translation of all spins in $k$ by a displacement $d\in\ZZ_L^D$ by
$k_d$, and define the translation-symmetrized basis tensors
\begin{align}
  \m m_k^{\t{TS}}
  \equiv \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \m m_{k_d}
  = \f1{\N_k^{\t{TS}}} \sum_{d\in\ZZ_L^D} \sum_{\sigma\in P_M}
  \m e_{\sigma(k_d)},
\end{align}
where the normalization factor $\N_k^{\t{TS}}$ enforces that the
entries of $\m m_k^{\t{TS}}$ are all either zero or one, which
otherwise might not occur if some translations of $k$ results in the
same choice of spins.  We then define the set $\T_{LD}\p{M}$ of
equivalence classes in $\C_N\p{M}$ under arbitrary translations on
$D$-dimensional lattice of linear size $L$, and expand
\begin{align}
  \m w = \sum_{k\in\T_{LD}\p{M}} w_k \m m_k^{\t{TS}}.
  \label{eq:sym_shift_tensor}
\end{align}
Having expanded $\m w$ in a basis that respects permutational symmetry
and translational invariance, we need to express the eigenvalue
equation \eqref{eq:H_VM_psi_eig} in the same basis.  In order to do
so, we label spins $p,q$ by the position of the corresponding lattice
site in $\ZZ_L^D$, and use translational invariance of the couplings
$\m h$ to expand
\begin{align}
  \m h \equiv \sum_{p,q\in\ZZ_L^D} h_{pq} \op{p}{q}
  = \sum_{q,d\in\ZZ_L^D} h_{q+d,q} \op{q+d}{q}
  = \sum_{d\in\ZZ_L^D} h_{d,0} \m s_d,
  &&
  \m s_d \equiv \sum_{q\in\ZZ_L^D} \op{q+d}{d}.
\end{align}
Starting from this form of $\m h$ and the expansion of $\m w$ in
\eqref{eq:sym_tensor}, we can work out that
\begin{align}
  \m h\circ\m w
  = \sum_{k\in\T_{LD}\p{M}} \sum_{d\in\ZZ_L^D} \sum_{a=1}^M w_{k_{d;a}}
  h_{d,0} w_{k_{d;a}} \m m_k^{\t{TS}},
\end{align}
where $k_{d;a}$ denotes the spins acquired by displacing the $a$-th
spin of $k$ by $d$.  With a bit of additional work, we can also
simplify
\begin{align}
  \bar{\m h}
  = \sum_{k\in\T_{LD}\p{M}} \sp{M h_0 - \sum_{p,q\in k} h_{pq}}
  \m m_k^{\t{TS}},
  &&
  h_0 \equiv \sum_{d\in\ZZ_L^D} h_{d,0},
\end{align}
which implies that
\begin{align}
  \bar{\m h} * \m w = M h_0 \m w
  - \sum_{k\in\T_{LD}\p{M}} \sum_{p,q\in k} h_{pq} w_k \m m_k^{\t{TS}}.
\end{align}
Putting everything together, the generalized eigenvalue problem in
\eqref{eq:H_VM_psi_eig} can be written in the form
\begin{align}
  \sum_{k\in\T_{LD}\p{M}} \sp{\sum_{p,q\in k} h_{pq} w_k
    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M w_{k_{d;a}}}
  \m m_k^{\t{TS}}
  = \p{M h_0 + \Delta} \m w,
\end{align}
or equivalently
\begin{align}
  \check{\m h} \c \v{\m w} = \p{M h_0 + \Delta} \v{\m w},
  \label{eq:H_VM_eig_simp}
\end{align}
where
\begin{align}
  \check{\m h} \equiv
  \sum_{k\in\T_{LD}\p{M}} \ket{k}
  \sp{\sum_{p,q\in k} h_{pq} \bra{k}
    + \sum_{d\in\ZZ_L^D} h_{d,0} \sum_{a=1}^M \bra{k_{d;a}}},
  &&
  \v{\m w} \equiv \sum_{k\in\T_{LD}\p{M}} w_k \ket{k},
\end{align}
and, to keep notation simple, we define $\ket{k}=0$ for all $k$ with
repeated values, and enforce $\ket{j}=\ket{k}$ if spins $j$ and spins
$k$ are the same up to a global translation.

@@

Weak $M$-body perturbations $\V_M$ to SU($n$)-symmetric interactions
can be treated perturbatively, resulting in an effective Hamiltonian
on a permutationally symmetric manifold $\M_0$ of $N$ spins.
Computing this effective Hamiltonian generally requires multiplying
several operators, and projecting the resulting product onto $\M_0$.
Even if the perturbation $\V_M$ is too strong for rigorous
perturbative treatment, as long as it is comparable in magnitude to
the energetic penalty incurred by breaking permutational symmetry, its
effects can still be accounted for by an effective theory on the joint
manifold of permutationally symmetric states together with states of
``weak'' permutational symmetry breaking, e.g.~the manifold of states
generated by the action of $M$-body perturbations on permutationally
symmetric states.  The expansion of operators in a basis for this
joint manifold likewise requires multiplying several operators and
projecting the resulting product onto $\M_0$.  Specifically, if
$\V_M=\sum_{k\in\C_N\p{M}}w_kO_k$, then we can define operators
\begin{align}
  \V_M^{(k)} \equiv \sum_{p\in\C_N\p{M}} w^{(k)}_p O_p,
\end{align}
where $\set{w^{(k)}}$ forms a complete basis for the set of symmetric
tensors of dimension $M$ with zeros on all diagonals (i.e.~zero in all
entries with a repeated index).  Denoting a complete basis of
permutationally symmetric states by $\set{\ket{m}}$, we can then
identify the $M$-body perturbed states
\begin{align}
  \ket{mk} \equiv \f1{\sqrt{\N_{mk}}} \V_M^{(k)} \ket{m},
  &&
  \N_{mk} \equiv \bk{m | \V_M^{(k)} \V_M^{(k)} | m}.
\end{align}
The restriction of a multi-body operator $\Q=\sum_p v_p Q_p$ to this
expanded Hilbert space is then
\begin{align}
  \Q =_M \P_M \Q \P_M,
  &&
  \P_M \equiv \sum_m \op{m} + \sum_{m,k} \op{mk},
\end{align}
where $=_M$ denotes equality up to a restriction to the support of
$\P_M$, and finding matrix elements of $\Q$ now requires computing
products of the form
\begin{align}
  \bk{\ell | \V_M^{(j)} \Q \V_M^{(k)} | m}
  = \sum_{p,q,r} w^{(j)}_p v_q w^{(k)}_r \bk{\ell | O_p Q_q O_r | m}.
  \label{eq:example_prod}
\end{align}
The purpose of this section is to work out such products in a
systematic way.  To this end, we derive a general diagrammatic
expansion of these products, as well as rules to simplify the
resulting expansions so as to minimize their computational cost.
These diagrammatic methods allow us to algorithmically evaluate
products such as \eqref{eq:example_prod} from the bare multi-body
operator data $\p{w^{(j)},O},\p{v,Q},\p{w^{(k)},O}$.  We then work out
some special cases of products that are relevant to the perturbation
theory in Sections \ref{sec:single_body_pert} and
\ref{sec:two_body_pert}.
